---
title: "Statistical Concepts"
---

# Assigment 2
Load the table random_vars.rds. You can either

* run the command: random_vars <- readRDS("your_download_folder/random_vars.rds")
* or find the file in the Files pane and select it.

After you load the table, get an overview of what it contains with View(random_vars).

* For each variable, compute the following values. You can use the built-in functions or use the mathematical formulas.
    + expected value
    + variance
    + standard deviation
* Explain, if it makes sense to compare the standard deviations.
* Then, examine the relationship between both variables and compute:
    + covariance
    + correlation
* What measure is easier to interpret? Please discuss your interpretation.
* Compute the conditional expected value for:
    + E[income | age <= 18]
    + E[income | age $\epsilon$ [18,65)]
    + E[income | age >= 65]
    
## Assignment 2.1
```{r}
#| include: false
#load libraries
library(dplyr)
```

```{r}
# Load the data from random_vars.rds file
data <- readRDS("data/random_vars.rds")  # Replace 'path_to_your_folder' with the actual path

# View the loaded data
View(data)

# Compute expected value, variance, and standard deviation for each variable
summary_stats <- data.frame(
  Variable = colnames(data),
  Expected_Value = sapply(data, mean),
  Variance = sapply(data, var),
  Standard_Deviation = sapply(data, sd)
)

# View summary statistics
summary_stats
```
## Assignment 2.2
Standard deviation provides insights into the variability of individual variables, comparing relationships between 'age' and 'income' would benefit more from measures like covariance and correlation to understand how these variables are related or associated with each other but comparing covariance is not easy. Where as in correlation, we get values between -1 and 1 which provides a standardized measure to interpret the strength and direction of linear relationship between age and income.

## Assignment 2.3
```{r}
# Compute covariance and correlation for each variable
summary_stats <- data.frame(
  Variable = colnames(data),
  Covariance = sapply(data, function(x) sapply(data, function(y) cov(x, y))),
  Correlation = sapply(data, function(x) sapply(data, function(y) cor(x, y)))
)

# View summary statistics
summary_stats
```
It is easier to interpret correlation as compared to covarience as discussed above.

## Assignment 2.4
```{r}
# Calculate conditional expected values for income based on age ranges
conditional_expected_value <- data %>%
  summarize(
    E_income_age_18 = mean(filter(., age <= 18)$income),
    E_income_age_18_65 = mean(filter(., age >= 18 & age < 65)$income),
    E_income_age_65 = mean(filter(., age >= 65)$income)
  )

conditional_expected_value
```

