---
title: "Statistical Concepts"
---

# Assigment 2
## Assignment 2.1
```{r}
#load libraries
library(dplyr)

# Load the data from random_vars.rds file
data <- readRDS("data/random_vars.rds")  # Replace 'path_to_your_folder' with the actual path

# View the loaded data
View(data)

# Compute expected value, variance, and standard deviation for each variable
summary_stats <- data.frame(
  Variable = colnames(data),
  Expected_Value = sapply(data, mean),
  Variance = sapply(data, var),
  Standard_Deviation = sapply(data, sd)
)

# View summary statistics
summary_stats
```
## Assignment 2.2
Standard deviation provides insights into the variability of individual variables, comparing relationships between 'age' and 'income' would benefit more from measures like covariance and correlation to understand how these variables are related or associated with each other but comparing covariance is not easy. Where as in correlation, we get values between -1 and 1 which provides a standardized measure to interpret the strength and direction of linear relationship between age and income.

## Assignment 2.3
```{r}
# Compute expected value, variance, and standard deviation for each variable
summary_stats <- data.frame(
  Variable = colnames(data),
  Covariance = sapply(data, function(x) sapply(data, function(y) cov(x, y))),
  Correlation = sapply(data, function(x) sapply(data, function(y) cor(x, y)))
)

# View summary statistics
summary_stats
```
It is easier to interpret correlation as compared to covarience as discussed above.

## Assignment 2.4
```{r}
# Calculate conditional expected values for income based on age ranges
conditional_expected_value <- data %>%
  summarize(
    E_income_age_18 = mean(filter(., age <= 18)$income),
    E_income_age_18_65 = mean(filter(., age >= 18 & age < 65)$income),
    E_income_age_65 = mean(filter(., age >= 65)$income)
  )

conditional_expected_value
```

