---
title: "Regression and Statistical Inference"
---

# Assignment 3

## Assignment 3.1
```{r}
#| include: false
#load libraries
library(dplyr)
library(ggplot2)
```

```{r}
#Read the data and check dimensions
data <- readRDS("data/car_prices.rds")
dim(data) # Check dimensions - rows and columns

```
## Assignment 3.2
```{r}
#Get a more detailed look at the data
str(data) # Display structure and data types
glimpse(data)
```
For strings <chr> Character and for numbers <dbl> Double data types are observed

## Assignment 3.3
```{r}
#We Run a linear regression 
model <- lm(price ~ ., data = data[, -20])
summary(model) # Check regression summary for coefficients, p-values, etc.
```
We want to now consider seeing p-values for selecting significant regressors affecting price and we construct a barplot with alpha set to 0.05 and shown with red dotted line. All those variables with p-value greater than 0.05 we don't consider them.

```{r}
#Extracting p-values of regressors from the summary of the regression model
p_values <- summary(model)$coefficients[, "Pr(>|t|)"]
p_values
# Extracting names of the regressors
regressor_names <- names(p_values)

# Create a bar plot to visualize the significance of regressors
barplot(-log10(p_values), 
        names.arg = regressor_names,
        main = "Significance of Regressors on Car Prices",
        xlab = "Regressors",
        ylab = "-log10(p-value)",
        col = "lightblue",
        ylim = c(0, max(-log10(p_values)) * 1.2),
        las = 2, # Rotate x-axis labels vertically
        cex.names = 0.7)

# Add a horizontal dashed line at significance level (for example, p = 0.05)
abline(h = -log10(0.05), lty = 2, col = "red")


# Determine significance levels based on p-values
significance_levels <- ifelse(p_values <= 0.001, "***",
                              ifelse(p_values <= 0.01, "**",
                                     ifelse(p_values <= 0.05, "*", "")))

# Filter significant regressors whose p-values are less than 0.05
significant_regressors <- regressor_names[p_values < 0.05]

# Create a table of significant regressors along with their significance levels
  significant_table <- data.frame(
  Regressors = significant_regressors,
  P_Values = p_values[p_values < 0.05],
  Significance = significance_levels[p_values < 0.05]
)
significant_table
# Include only significant regressors
model_imp <- lm(price ~ enginelocation + enginesize + stroke + carwidth + peakrpm, data = data)
summary(model_imp)

# Extract the coefficients of significant regressors from the regression model summary
significant_regressors <- c("enginelocationrear", "enginesize", "stroke", "carwidth", "peakrpm")

# Extract the coefficients and variable names
coefficients <- coef(model)[significant_regressors]

# Create a bar plot to visualize the impact of significant regressors on price
barplot(coefficients, 
        main = "Impact of Significant Regressors on Car Prices",
        xlab = "Regressors",
        ylab = "Coefficient Estimate",
        col = "skyblue",
        ylim = c(min(coefficients) * 1.2, max(coefficients) * 1.2),
        las = 2, # Rotate x-axis labels vertically
        beside = TRUE)

# Add labels for the variables
text(1:length(coefficients), coefficients, labels = round(coefficients, digits = 2), pos = 3, cex = 0.8)
```

## Assignment 3.4
Now for this part enginesize is selected based on its p-value (4.887489e-06) and significance and also a positive correletion based on the plot and positive estimate of 125.72 show that as enginesize increases, price also increases. The data type of enginesize is <dbl> double and it is statistically significant since its p-value is less than alpha= 0.05.  
```{r}
# Plot relationship between enginesize and price
ggplot(data, aes(x = enginesize, y = price)) +
  geom_point(alpha = 0.8)

```
```{r}
# Step 5: Add a new variable 'seat_heating' and run regression
data_with_seat_heating <- data %>% mutate(seat_heating = TRUE)
model_with_seat_heating <- lm(price ~ ., data = data_with_seat_heating)
summary(model_with_seat_heating) # Check regression summary for the coefficient of seat_heating

```
Adding a variable seat_heating to the data and assigning a value TRUE for all observations gave a NA for Estimate coefficient in the new regression model. This new variable has no affect on the price since it is same for all observations.
