---
title: "Directed Acyclic Graphs"
---
# Assignment 5

* Think about example from previous chapter (parking spots) and draw the DAG.

* Load the data file customer_sat.rds. A company sells a SaaS (Software as a service) to business clients that chose to subscribe to one of the three subscription levels, Starter, Premium or Elite, which provide different features at different price points (more storage and other capabilities). Starter has the least functionality, while Elite includes the most features.

  Now imagine, you are working as an analyst for that company and are asked to examine the relationship between follow-up calls to the clients and their satisfaction with the product and service. Follow-up calls are voluntary and are just to help the client with possible problems and so on. Satisfaction is measured on a weekly basis by surveying the clients.

  In the data, you find three variables: satisfaction, follow_ups and subscription. Perform the following steps:

    + regress satisfaction on follow_ups

    + regress satisfaction on follow_ups and account for subscription

* Compare the coefficients and find a possible explanation.

* Plot the data. You’ll find a similar plot in this chapter, that you can copy some lines from.

```{r}
#| include: false
# Load necessary libraries
library(dagitty)
library(ggdag)
library(tidyverse)
```
## Assignment 5.1
```{r}
# Define the DAG for the parking spot example
parking_dag <- dagify(
  sales ~ parking_spots,
  sales ~ location,
  parking_spots ~ location,
  coords = list(x = c(sales = 1, parking_spots = 3, location = 2),
                y = c(sales = 0, parking_spots = 0, location = 1))
)


# Plot the DAG with default ggplot2 theme
ggdag(parking_dag, text = FALSE) +
  geom_dag_point() +
  geom_dag_text(color = "red") +
  geom_dag_edges(edge_color = "black")

```
## Assignment 5.2.1
```{r}
# Load the data file customer_sat.rds
data <- readRDS("data/customer_sat.rds")

# 1. Regress satisfaction on follow_ups
model_1 <- lm(satisfaction ~ follow_ups, data = data)

summary(model_1)
```
## Assignment 5.2.2
```{r}
# 2. Regress satisfaction on follow_ups and account for subscription
model_2 <- lm(satisfaction ~ follow_ups + subscription, data = data)

summary(model_2)
```
## Assignment 5.3
According to the first regression, there is a negative relationship between follow_ups and satisfaction, indicating that as the number of follow-up calls increases, so does satisfaction.

The second regression, which takes into account both follow_ups and subscription levels, reveals that, while follow_ups continue to have an impact on satisfaction, subscription levels (Premium and Premium+) also have a significant impact. Furthermore, the subscription level coefficients imply that higher subscription tiers (Premium and Premium+) have a positive impact on satisfaction when compared to some assumed base level (possibly Starter, which is not explicitly included as a variable).

## Assignment 5.4
```{r}
# Create a DAG for the relationship between follow-up calls, satisfaction, and subscription
saas_dag <- dagify(
  satisfaction ~ follow_ups,
  satisfaction ~ subscription,
  follow_ups ~ subscription,
  coords = list(x = c(satisfaction = 1, follow_ups = 3, subscription = 2),
                y = c(satisfaction = 0, follow_ups = 0, subscription = 1))
)

# Plot the DAG 
ggdag(saas_dag, text = FALSE) +
  geom_dag_point() +
  geom_dag_text(color = "red") +
  geom_dag_edges(edge_color = "black")



```