---
title: "Randomized Controlled Trials"
---

# Assignment 6

Load the dataset abtest_online.rds. It contains data about a randomized experiment run by an online shop. E-commerce websites frequently conduct numerous randomized experiments, commonly referred to as AB testing in a business context. In these experiments, a subset of randomly chosen website visitors is presented with a slightly altered version of the site, while others see the standard version. This approach enables the testing of new features, and the decision to implement a new feature is contingent on the outcomes derived from these tests.

Consider this situation: You operate an online store and are concerned about the expenses associated with your customer service. To cut costs, you contemplate introducing a chatbot as a substitute for human customer service. However, you’re uncertain whether this change might have a detrimental impact on your sales. Consequently, you intend to conduct an AB test. In this test, a portion of users will be directed to a website equipped with a chatbot (treatment group), while the remaining customers will still interact with human customer service (control group) if they have questions. To ensure randomization, you’ll allocate the treatment based on the last digit of each user’s IP address.

There are two outcome variables, purchase and purchase_amount. The first one shows whether a customer bought and the other how much (in €) he bought. First, let’s use purchase_amount.

Other variables included are mobile_device being TRUE when a user visits the site using a mobile device and previous_visits indicating the number of previous visits of a particular user.

After loading the data, perform the following steps:

  * Check whether the covariates are balanced across the groups. Use a plot to show it.
  * Run a regression to find the effect of chatbot on sales.
  * Find subgroup-specific effects by including an interaction. Compute a CATE for one exemplary group. A subgroup    could be for example mobile users.
  * It’s not only of interest how much customers buy but also if the buy at all. Then, the dependent variable is binary (either 0 or 1) instead of continuous and the model of choice is the logistic regression. Use the outcome variable purchase and run a logistic regression. The coefficients are not as easily interpretable as before. Look it up and interpret the coefficient for chatbot.

```{r}
#| include: false
#load libraries
library(dplyr)
library(ggplot2)
```

## Assignment 6.1
```{r}
# Loading dataset into 'data'
data <- readRDS("data/abtest_online.rds")
str(data)

# Convert non-numeric variables to numeric
data$mobile_device <- as.numeric(data$mobile_device)
data$chatbot <- as.numeric(data$chatbot)

# Check covariate balance across groups using a plot 
covariates <- c("mobile_device", "previous_visit", "purchase", "purchase_amount") # Variables of interest
group_var <- "chatbot"

# Check covariate balance visually using plots
plot_balances <- lapply(covariates, function(var) {
  ggplot(data, aes_string(x = group_var, y = var, color = as.factor(group_var))) +
    stat_summary(
      geom = "errorbar", 
      width = 0.5,
      fun.data = "mean_se", 
      fun.args = list(mult = 1.96),
      show.legend = FALSE
    ) +
    labs(title = paste("Covariate Balance for", var))
})
plot_balances
```
Plots shows in general, covariates are overall balanced when chatbot is True and False. However, when comparing previous_visits plots it can be seen that the CI(Confidence Intervals) overlap with each other.


## Assignment 6.2: Running a regression on 'purchase_amount' to find the effect of chatbot on sales
```{r}
sales_model <- lm(purchase_amount ~ chatbot, data = data)
summary(sales_model) # regression summary
```
The summary shows that the sales dropped for customers that used the chatbot with the chatbot coefficient value = -7.0756 

## Assignment 6.3: Interaction between 'chatbot' and 'mobile_device' for subgroup effects
```{r}
interaction_model <- lm(purchase_amount ~ chatbot * mobile_device, data = data)
summary(interaction_model) # regression summary (-0.1526 for chatbot:mobile_device)

# Computing CATE for mobile users (subgroup)
mobile_users <- data[data$mobile_device == 1, ]

# Fit a model for mobile users
mobile_users_model <- lm(purchase_amount ~ chatbot, data = mobile_users)
summary(mobile_users_model) # View regression summary
```

## Assignment 6.4: Run logistic regression for binary outcome 'purchase'
```{r}
logistic_model <- glm(purchase ~ chatbot, 
                      family = binomial(link = 'logit'), data = data)
summary(logistic_model) # View regression summary

# Now Interpreting the coefficient for 'chatbot'
# Coefficients here represent log odds ratio. In order to get the odds ratio, coefficient for chatbot is exponentiated.
exp(coef(logistic_model)["chatbot"]) # Exponentiated coefficient for 'chatbot'
```
The odds ratio amount to 0.371 which suggests that the customers are less likely to purchase something with the chatbot

```


