---
title: "Randomized Controlled Trials"
---

# Assignment 6
```{r}
library(dplyr)
library(ggplot2)
```

## Assignment 6.1
```{r}
# Loading dataset into 'data'
data <- readRDS("data/abtest_online.rds")
str(data)

# Convert non-numeric variables to numeric
data$mobile_device <- as.numeric(data$mobile_device)
data$chatbot <- as.numeric(data$chatbot)

# Check covariate balance across groups using a plot 
covariates <- c("mobile_device", "previous_visit", "purchase", "purchase_amount") # Variables of interest
group_var <- "chatbot"

# Check covariate balance visually using plots
plot_balances <- lapply(covariates, function(var) {
  ggplot(data, aes_string(x = group_var, y = var, color = as.factor(group_var))) +
    stat_summary(
      geom = "errorbar", 
      width = 0.5,
      fun.data = "mean_se", 
      fun.args = list(mult = 1.96),
      show.legend = FALSE
    ) +
    labs(title = paste("Covariate Balance for", var))
})
plot_balances
```
Plots shows in general, covariates are overall balanced when chatbot is True and False. However, when comparing previous_visits plots it can be seen that the CI(Confidence Intervals) overlap with each other.


## Assignment 6.2: Running a regression on 'purchase_amount' to find the effect of chatbot on sales
```{r}
sales_model <- lm(purchase_amount ~ chatbot, data = data)
summary(sales_model) # regression summary
```
The summary shows that the sales dropped for customers that used the chatbot with the chatbot coefficient value = -7.0756 

## Assignment 6.3: Interaction between 'chatbot' and 'mobile_device' for subgroup effects
```{r}
interaction_model <- lm(purchase_amount ~ chatbot * mobile_device, data = data)
summary(interaction_model) # regression summary (-0.1526 for chatbot:mobile_device)

# Computing CATE for mobile users (subgroup)
mobile_users <- data[data$mobile_device == 1, ]

# Fit a model for mobile users
mobile_users_model <- lm(purchase_amount ~ chatbot, data = mobile_users)
summary(mobile_users_model) # View regression summary
```

## Assignment 6.4: Run logistic regression for binary outcome 'purchase'
```{r}
logistic_model <- glm(purchase ~ chatbot, 
                      family = binomial(link = 'logit'), data = data)
summary(logistic_model) # View regression summary

# Now Interpreting the coefficient for 'chatbot'
# Coefficients here represent log odds ratio. In order to get the odds ratio, coefficient for chatbot is exponentiated.
exp(coef(logistic_model)["chatbot"]) # Exponentiated coefficient for 'chatbot'
```
The odds ratio amount to 0.371 which suggests that the customers are less likely to purchase something with the chatbot

```


