---
title: "Difference-in-Differences"
---

# Assignment 8

Imagine, you are manager of a large health provider that manages many hospitals and you want to test how a new admission procedure affects patient satisfaction. You randomly selected 18 hospitals that introduced the new admission procedure and compare them to 28 other hospitals that did not introduce the method. For both groups of hospitals you collected data from before and after the introduction. The data you have collected is from patient surveys where they were asked how satisfied they are.

Load the data from the file hospdd.rds1. Then, perform a difference-in-differences analysis by

   * Manually computing the mean satisfaction for treated and control hospitals before and after the treatment. Helpful functions could be filter(), pull() and basic arithmetic operations.

   * Using a linear regression to compute the estimate. Also, include group and time fixed effects in the regression, i.e. one regressor for each month and one regressor for each hospital: Consider, whether you want to include them as

      + month + hospital or as

      + as.factor(month) + as.factor(hospital)

  and explain what the difference is.

## Assignment 8.1
```{r}
#load libraries
library(Matching)
library(ggdag)
library(tidyverse)
library(dagitty)
library(MatchIt)
```

```{r}
# loading data into data
data <- readRDS("data/hospdd.rds")
data
names(data)
mean_before_treatment <- data %>%
  filter(procedure == 0, hospital>= 1 & hospital<=18) %>%
  pull(satis) %>%
  mean()

mean_after_treatment <- data %>%
  filter(procedure == 1, hospital>= 1 & hospital<=18) %>%
  pull(satis) %>%
  mean()

mean_before_control <- data %>%
  filter(procedure == 0, hospital>18 & hospital<= 46) %>%
  pull(satis) %>%
  mean()

mean_after_control <- data %>%
  filter(procedure == 0, hospital>18 & hospital<= 46) %>%
  pull(satis) %>%
  mean()
diff_before <- mean_before_treatment - mean_before_control
diff_after <- mean_after_treatment - mean_after_control

diff_in_diff <- diff_after - diff_before

sprintf("Mean satisfaction for treated hospitals before treatment: %f" ,mean_before_treatment)
sprintf("Mean satisfaction for treated hospitals after treatment: %f", mean_after_treatment)
sprintf("Mean satisfaction for control hospitals before treatment: %f" , mean_before_control)
sprintf("Mean satisfaction for control hospitals after treatment: %f", mean_after_control)
sprintf("Difference in difference: %f", diff_in_diff)


```
## Assignment 8.2
```{r} 

# Linear regression with month and hospital as continuous variables
model_continuous <- lm(satis ~ month + hospital + procedure, data = data)
summary(model_continuous)


# Linear regression with factors for month and hospital
model_factors <- lm(satis ~ as.factor(month) + as.factor(hospital) + procedure, data = data)
summary(model_factors)
```
The model using categorical variables for 'month' and 'hospital' demonstrates a better fit explaining 52.99% of the variance in patient satisfaction and improved explanatory power compared to the continuous variable model. Individual hospitals and certain months have significant impacts on patient satisfaction besides the effect of the procedure.
