---
title: "Difference-in-Differences"
---

# Assignment 8

## Assignment 8.1
```{r}
#load libraries
library(Matching)
library(ggdag)
library(tidyverse)
library(dagitty)
library(MatchIt)
```

```{r}
# loading data into data
data <- readRDS("data/hospdd.rds")
data
names(data)
mean_before_treatment <- data %>%
  filter(procedure == 0, hospital>= 1 & hospital<=18) %>%
  pull(satis) %>%
  mean()

mean_after_treatment <- data %>%
  filter(procedure == 1, hospital>= 1 & hospital<=18) %>%
  pull(satis) %>%
  mean()

mean_before_control <- data %>%
  filter(procedure == 0, hospital>18 & hospital<= 46) %>%
  pull(satis) %>%
  mean()

mean_after_control <- data %>%
  filter(procedure == 0, hospital>18 & hospital<= 46) %>%
  pull(satis) %>%
  mean()
diff_before <- mean_before_treatment - mean_before_control
diff_after <- mean_after_treatment - mean_after_control

diff_in_diff <- diff_after - diff_before

sprintf("Mean satisfaction for treated hospitals before treatment: %f" ,mean_before_treatment)
sprintf("Mean satisfaction for treated hospitals after treatment: %f", mean_after_treatment)
sprintf("Mean satisfaction for control hospitals before treatment: %f" , mean_before_control)
sprintf("Mean satisfaction for control hospitals after treatment: %f", mean_after_control)
sprintf("Difference in difference: %f", diff_in_diff)


```
## Assignment 8.2
```{r} 

# Linear regression with month and hospital as continuous variables
model_continuous <- lm(satis ~ month + hospital + procedure, data = data)
summary(model_continuous)


# Linear regression with factors for month and hospital
model_factors <- lm(satis ~ as.factor(month) + as.factor(hospital) + procedure, data = data)
summary(model_factors)
```
The model using categorical variables for 'month' and 'hospital' demonstrates a better fit explaining 52.99% of the variance in patient satisfaction and improved explanatory power compared to the continuous variable model. Individual hospitals and certain months have significant impacts on patient satisfaction besides the effect of the procedure.
